---
title: "AmpliPub analysis report"
output:
  html_document:
    toc: true
    toc_float: true
    self_contained: true
params:
  results: ""
  run_info: ""
  provenance_dir: ""
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
res <- readRDS(params$results)
out_dir <- dirname(params$results)
# Figures are written into the HTML as base64 data by knitr::image_uri. Linking them with
# include_graphics left relative paths that pandoc could not embed from the temporary knit
# directory, and inside a loop the linked figures did not appear at all. Every figure chunk
# is results = "asis".
fig <- function(name) {
  path <- file.path(out_dir, "figures", paste0(name, ".png"))
  if (file.exists(path)) {
    cat(sprintf('\n<p><img src="%s" alt="%s" style="max-width:100%%;"></p>\n',
                knitr::image_uri(path), name))
  } else {
    cat("\nFigure `", name, "` was not drawn; see the analysis log.\n", sep = "")
  }
  invisible(NULL)
}
show <- function(obj) cat(paste(utils::capture.output(print(obj)), collapse = "\n"))
tbl <- function(df, digits = 4) knitr::kable(df, digits = digits, row.names = FALSE)
```

This report reads the outputs of the `amplipub_analysis` workflow stage and computes nothing
new. Every number here is also in `tables/`, and every figure in `figures/`.

# Run

```{r run-info}
tbl(utils::read.delim(params$run_info, stringsAsFactors = FALSE))
```

```{r git, results = "asis"}
commit <- file.path(params$provenance_dir, "git_commit.txt")
if (file.exists(commit)) cat("Repository commit: `", readLines(commit, warn = FALSE)[1], "`\n\n", sep = "")
status <- file.path(params$provenance_dir, "git_status.txt")
if (file.exists(status) && length(readLines(status, warn = FALSE)) > 0L) {
  cat("**The checkout had uncommitted changes when this ran.** See `provenance/git_status.txt`.\n")
}
```

# Study design

```{r scan, results = "asis"}
cat("```\n"); show(res$scan); cat("\n```\n")
```

# Alpha diversity

Computed on the rarefied table from `10_diversity` (depth `r res$depth`).

```{r alpha-tests}
tbl(res$alpha_test$results)
```

```{r alpha-fig, results = "asis"}
fig("alpha")
```

# Beta diversity

```{r permanova}
tbl(merge(res$permanova$results, res$permanova$interpretation[, c("metric", "term", "verdict")],
          by = c("metric", "term")))
```

```{r permanova-text, results = "asis"}
for (i in seq_len(nrow(res$permanova$interpretation))) {
  r <- res$permanova$interpretation[i, ]
  cat("- **", r$metric, "**: ", r$interpretation, "\n", sep = "")
}
```

```{r ordination-figs, results = "asis"}
for (m in res$beta$metrics) {
  cat("\n\n## ", m, "\n\n", sep = "")
  fig(paste0("ordination_", m))
  fig(paste0("dispersion_", m))
}
```

# Composition

```{r taxa-figs, results = "asis"}
fig("taxa_phylum_bars")
fig("taxa_genus_bars")
fig("taxa_genus_heatmap")
```

# Differential abundance

```{r da, results = "asis"}
cat("```\n"); show(res$concordance); cat("\n```\n")
```

```{r da-figs, results = "asis"}
fig("da_volcano")
fig("da_concordance")
fig("da_effects")
```

# Exploratory screen

Hypothesis-generating. Ranked by variance explained adjusted for degrees of freedom, with one
multiple-testing correction across every test run.

```{r screen}
tbl(utils::head(res$screen$results[, c("rank", "variable", "family", "metric", "effect_name",
                                       "effect_adj", "p", "q", "stability", "verdict")], 30))
```

```{r screen-fig, results = "asis"}
fig("screen")
```

```{r explains, results = "asis"}
if (!is.null(res$explains)) {
  cat("\n# Confirmatory model\n\n")
  cat("```\n"); show(res$explains); cat("\n```\n")
}
```

# Normalization sensitivity

Whether each conclusion holds under every normalization compared.

```{r normalization}
tbl(res$normalization$results)
tbl(res$normalization$agreement)
```

# Software

```{r session, results = "asis"}
cat("```\n", paste(readLines(file.path(out_dir, "provenance", "sessionInfo.txt")), collapse = "\n"), "\n```\n")
envs <- list.files(params$provenance_dir, pattern = "^env_.*\\.yml$")
if (length(envs)) cat("\nConda environment exports: ", paste0("`", envs, "`", collapse = ", "), "\n")
```
